---
title: "STAT 27410 Final Project Proposal"
author: "Ariana Qin & Robert Rodriguez"
fontsize: 12pt
bibliography: references.bib
output:
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
  word_document: default
  geometry: margin=0.75in
fig_crop: no
header-includes:
   - \usepackage{booktabs}
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F)
options(scipen=6, digits=6)
spks = read_csv("spks.csv")
redlabels = read_csv("redlabels.csv")
cell_pos = read_csv("cell_pos.csv")
calcium_reach_times = read_csv("calcium_reach_times.csv")
```

# 1. Introduction

Complex motor skills, which are precise and non-stereotypical movements, are essential for survival. For example, squirrels need to jump from branch to branch, and eagles need to catch small prey. Understanding brain activity during complex motor skills is essential in motor control and motor learning. Towards this goal, the MacLean Lab, a neuroscience lab at The University of Chicago, performed neuro-recording on the mouse primary motor (M1) cortex while the mouse was engaged in a reach-to-grasp movement. The M1 cortex can be seen as a recurrent neural network, receiving upstream signals on motor planning and body position, and outputting commands to downstream areas, such as the spinal cord and brainstem. While networks are very complex objects of study, we chose a narrow objective for this project: for any neuron in the network, we want to predict its activity using the activity of other neurons. We hope this can inspire future research on the dynamics and structure of neural networks.

## 1.1 Dataset

We use the data from the MacLean lab. During the experiment, the mouse reached for a rice grain in a head-fixed position. Before the experiment, the mouse underwent surgery so that its brain is attached to a microscope and that neurons would increase in fluorescence value when their activity increased. After the initial imaging, the MacLean Lab performed preprocessing to infer the estimated spike counts from the fluorescence data. 

\begin{itemize}
\item Time series data of neural spikes: This is a time series of the inferred number of spikes at each time bin, inferred from the fluorescence values recorded by the microscope. Each time bin is 30 ms. We have 315 neurons and 148468 time bins.

\item Inhibitory or excitatory neurons: This is an array that stores whether a neuron is inhibitory or excitatory, namely whether it increases or decreases the level of activity in downstream neurons.

\item Spatial distribution: This is an array of 2-tuples that stores the x, y coordinate of each neuron.

\item Mouse behavior: This is an array of 2-tuples that stores the start and end time of each bout of reaches.
\end{itemize}

## 1.2 Complications of Dataset

The dataset chosen is not without complications, however, some of these complications include:

\begin{itemize}
\item Incomplete network: Due to limitations of imaging techniques, the experiment records an incomplete network. We cannot record the complete M1 cortex, and not even all the neurons in the field of view. Hence, the recording is a small subset of neurons in the M1 cortex. 

\item Upstream signal: This entire brain region receives shared upstream signals. However, we cannot measure the upstream signals.
\end{itemize}

## 1.3 Scientific Question

For a target neuron, the goal is to predict its activity at each time bin using the activity of other neurons. In other words, the response variable is the activity of the target neuron. The predictor variables are the activity of the other neurons. The selected target neurons are Neuron 234, 168, and 281. The choice of selecting these three neurons comes from exploratory data analysis which showed that these three neurons had the potential to create relatively strong predictive models, making them neurons of interest and the three that we believe will provide the most interesting exploration and analysis of our data.

It is known that the activity of a target neuron can be explained by the activity of other neurons and their recurrent connections through linear models (Kotekal et al., 2020). This inspired us to construct linear regression models to capture the interaction between the predicted and predictor variables. In the frequentist analysis section, we hypothesize that different neurons integrate the activities of other neurons differently. This is a plausible hypothesis because the different neurons likely form different connections with other neurons, depending on factors such as their proximity. We seek to validate this hypothesis by comparing linear models that capture different integrations and linear models that require the same integration. This hypothesis, if true, would motivate hierarchical models in our Bayesian analysis. 

Due to the high computational cost, we cannot perform the complete frequentist and Bayesian analysis on all 295 excitatory neurons. Instead, we pick three neurons. Because the experiment only records an incomplete network, the selection criterion is based on the information available regarding a neuron’s upstream neurons. (See further explanation in the frequentist analysis section.)

# 2. Exploratory Data Analysis

For the exploratory analysis, we are interested in visualizing the neuron activity and co-activity, neuron spatial distribution, type (inhibitory or excitatory), and mouse behavior. 

## 2.1 Neuron Activity:

```{r figs, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs}Time Series Plot of Neuron 111's Activity"}
spks_t = as.data.frame(t(spks))
neurons = paste("Neuron", 1:315, sep = "_")
colnames(spks_t) = neurons

time_points = seq(0, by = 30 / 1000, length.out = 148468)
spks_t$time = time_points

ggplot(spks_t, aes(x = time, y = Neuron_234)) +
  geom_line(size = 1) +
  labs(title = "Time Series", x = "Time", y = "Inferred Number of Spikes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

As an example, we visualize the spikes time series of one of our target neurons, neuron 234 (Fig. 1). At each time bin, the y-axis displays the inferred number of spikes. A common question is why the quantity is continuous rather than discrete. Recall that spikes are not directly observable from the experiment. Instead, they are inferred from the observed fluorescence values of the neurons. Hence, the inferred number of spikes results from the spikes inference algorithm. We recommend interpreting the y-axis as an artificial measure of neuron activity level, and we refer to this quantity as inferred spikes. For more details, read the frequently asked questions section of the spike inference algorithm: https://github.com/HelmchenLabSoftware/Cascade. 


## 2.2 Neuron Types:

```{r figs2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs2}Distribution of Excitatory vs Inhibitory Neurons"}
colnames(redlabels) = c("neuron_type")
redlabels$neuron_type = factor(redlabels$neuron_type, levels= c(TRUE, FALSE), labels=c("Inhibitory", "Excitatory"))
ggplot(redlabels, aes(x = neuron_type)) +
  geom_bar() +
  theme_minimal() +
  labs(title= "Number of Excitatory and Inhibitory Neurons", x = "Neuron Type", y = "Number of Neurons") +
  theme(plot.title = element_text(hjust = 0.5))
```

Fig. 2 displays the counts of inhibitory and excitatory neurons in a barplot. Inhibitory and excitatory neurons are two types neurons, classified based on how they interact with downstream neurons. Each neuron has a membrane potential, namely the electric potential difference between the inside and outside of the cell. When a neuron's membrane potential reaches a certain threshold, it generates a spike, namely a sudden increase in action potential. This spike transmit signals to its downstream neurons and changes the action potential of the downstream neurons. Based on whether a neuron increases or decreases the action potential of its downstream neuron, the neuron is classified as excitatory or inhibitory respectively. The bar plot shows that there are many more excitatory neurons than inhibitory neurons. 

## 2.3 Correlation between Neuron Spikes:

```{r figs3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs3}Heatmap of Pearson Correlation of Neuron Spikes"}
spks_t = as.data.frame(t(spks))
cor_matrix = cor(spks_t, use = "pairwise.complete.obs")
diag(cor_matrix) = NA

library(pheatmap)
pheatmap(cor_matrix, 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         main = "Pearson Correlation Heatmap",
         treeheight_row = 0,
         treeheight_col = 0,
         show_rownames = FALSE,
         show_colnames = FALSE
)
```

In addition to single-neuron activity, we are interested in the collective dynamics of neurons. Here, we examine the co-activity between every pair of neurons, measured by Pearson correlation. The resulting matrix is a symmetric 315 by 315 matrix, where the ij-th entry is the Pearson correlation between the inferred spikes of the ith neuron and the jth neuron. Fig. 3 visualizes the Pearson correlation matrix using a heatmap. The heatmap shows that the correlation between neuron activity is low across the board. In addition, most correlations are positive, and negative correlations are small in magnitude. 

## 2.4 Neuron Positions:

```{r figs4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs4}Spacial Positioning of Neurons"}
colnames(cell_pos) = c("x", "y")
color_pos = cbind(cell_pos, redlabels)
ggplot(color_pos, aes(x = x, y = y, color = as.factor(neuron_type))) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("blue", "red")) +
  labs(title="Spatial Position of Neurons", color = "Neuron Type") +
  theme(plot.title = element_text(hjust = 0.5))
```

Fig 4. displays the spatial positions of the 315 excitatory and inhibitory neurons. x, y are the coordinates of the neurons in the image. Observe that some areas have more recorded neurons than others. In addition, because of the biophysics constraints, the neurons in proximity are more likely to be connected. Therefore, we expect the neurons in proximity to be more correlated. 

## 2.5 Reach Times:

```{r figs5, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs5}Distribution of Bout Lengths"}
colnames(calcium_reach_times) = c("start", "end")
calcium_reach_times$reach_time = calcium_reach_times$end - calcium_reach_times$start
ggplot(calcium_reach_times, aes(x=reach_time)) +
  geom_histogram(binwidth=20) +
  theme_minimal() +
  labs(title= "Distribution of Bout Lengths", x = "Bout Lengths", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

In the experiment, the mouse performs a reach-to-grasp task. Specifically, the mouse reaches for a tiny grain of rice. Given the task's difficulty, the mouse often makes several continuous attempts between rests. We call a set of continuous attempts “a bout of reaches”. Fig. 5 displays a histogram of the lengths of the bouts, given by subtracting the start time of the bout from the end time of the bout. We see that most bouts are short, but there are a few long bouts of reaches.

# 3. Frequentist Analysis

## 3.1 Proposed Frequentist Model(s)

Our project aims to predict the inferred spikes of a target neuron using the activity of the other neurons. In the frequentist paradigm, our model of choice is linear regression. Linear models are a well-established method in computational neuroscience literature. Specifically, Kotekal et al. 2020 showed that recurrent interactions (i.e., activities of other neurons) can explain target neuron activity through a linear model.

In addition to past literature, our choice of linear regression holds several other advantages. Because of its relative simplicity, it has a closed-form solution and is computationally efficient. Given our model's large number of predictors, computational complexity is an important consideration. Another advantage of simplicity is the interpretability of linear regression. We can understand how the predictors contribute to the target neuron activity and interpret the parameters as the strength of interaction between the two neurons. Given linear models' computational efficiency and interpretability, it is our model of choice. 

Using linear regression, we predict the inferred spikes of a target neuron using the inferred spikes of other neurons. To measure the performance of linear regression, we use the R^2 value, which is the percent of variance explained. In neuroscience, an R^2 around 0.1 is considered a good model. 

Due to the computational cost of performing the complete frequentist and Bayesian analysis for all 295 excitatory neurons, we pick 3 neurons to analyze for this project. The criterion for neuron choice is as follows: Because our data represents an incomplete network, we want to pick neurons whose upstream neurons are observable (i.e., included in the data). Otherwise, we do not have information on the predictors, and the subsequent analysis wouldn’t be meaningful. To determine the neurons with observable upstream neurons, we first perform a linear regression on all 295 neurons. We take the 3 neurons with the highest R^2 value.  Fig. 6 displays a histogram of the R^2 of all the neurons. 

```{r figs6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs6}Histogram of R^2 Values for Regression Models"}
load("r_squared_list.RData")
r_squared_vector <- unlist(r_squared_list)

hist(r_squared_vector, 
     main = "Histogram of R-squared values", 
     xlab = "R-squared", 
     col = "skyblue", 
     border = "black")
```

Using this method, we pick neurons 234, 168, and 281. 

For these three neurons, we consider two models for comparison. 

For these three neurons, we consider two models for comparison.

Model 1: 

$$ y_{\text{234}} = \alpha_{\text{234}} + X \beta_{\text{234}} $$

$$ y_{\text{168}} = \alpha_{\text{168}} + X \beta_{\text{168}} $$

$$ y_{\text{281}} = \alpha_{\text{281}} + X \beta_{\text{281}} $$


Model 2: 

$$ \tilde{y} =\mathbf{1} \tilde{\alpha} + \tilde{X} \tilde{\beta}$$

Where: 

Model 1: 

\begin{itemize}
\item$y_n$ is the inferred spikes of neuron n, an array of dimension (number of time bins) by 1. 

\item$X$ stores the time series of all neurons other than neurons 234, 168, and 281. It is a matrix of dimension (number of time bins) by (number of other neurons)

\item$\alpha_n$ and $\beta_n$ are parameters to be fitted. 
\end{itemize}

Model 2:
\begin{itemize}
\item$ \tilde{y}$ is the time series of neurons 234, 168, and 281, concatenated vertically. It is an array of dimension ( $3 \times \text{number of time bins}$) by 1

\item$ \tilde{X}$ is the time series of all neurons other than neurons 234, 168, and 281, concatenated three times vertically. It is a matrix of dimension ( $3 \times \text{number of time bins}$) by (number of other neurons).

\item$\tilde{\alpha}$ and $\tilde{\beta}$ are parameters to be fitted. 
\end{itemize}

We chose these two models to compare because we want to understand if neurons integrate the activities from other neurons similarly or differently. In model 1, the coefficient for the predictors could be different across neurons 234, 168, and 281, representing different integrations of the activity of other neurons. On the other hand, model 2 requires the coefficient and integration to be the same. Based on neuroscience knowledge, we hypothesize that model 1 will have better performance, as different neurons have different connectivity with other neurons and integrate signals differently. 

## 3.2 Fitting the Frequentist Model(s)

The two frequentist models were fit by compiling the predictor neurons into matrices and solving for Beta using the following equation:

$$
\hat{\beta} = (X^T X)^{-1} (X^T y)
$$

The resulting Beta is a column vector containing the intercept value and the 312 coefficients for the predictors. The models for Neurons 234, 168, and 281 are representative of our first frequentist model, which creates 3 separate models for each Neuron that we are analyzing and trying to predict for. The concatenated model is representative of our second frequentist model, which compiles all three response variables into a single matrix and tries to predict for all of them together.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
spks_t = as.data.frame(t(spks))
neurons = paste("Neuron", 1:315, sep = "_")
colnames(spks_t) = neurons

response_neurons = c("Neuron_234", "Neuron_168", "Neuron_281")
predictor_neurons = setdiff(neurons, response_neurons)
X = as.matrix(spks_t[, predictor_neurons])
X = cbind(Intercept = 1, X)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
Y1 = spks_t[, "Neuron_234"]

beta1 = solve(t(X) %*% X) %*% t(X) %*% Y1
Y1_pred = X %*% beta1
residuals1 = Y1 - Y1_pred

RSS1 = sum((Y1 - Y1_pred)^2)
TSS1 = sum((Y1 - mean(Y1))^2)
R_squared1 = 1 - RSS1 / TSS1

Y2 = spks_t[, "Neuron_168"]

beta2 = solve(t(X) %*% X) %*% t(X) %*% Y2
Y2_pred = X %*% beta2
residuals2 = Y2 - Y2_pred

RSS2 = sum((Y2 - Y2_pred)^2)
TSS2 = sum((Y2 - mean(Y2))^2)
R_squared2 = 1 - RSS2 / TSS2

Y3 = spks_t[, "Neuron_281"]

beta3 = solve(t(X) %*% X) %*% t(X) %*% Y3
Y3_pred = X %*% beta3
residuals3 = Y3 - Y3_pred

RSS3 = sum((Y3 - Y3_pred)^2)
TSS3 = sum((Y3 - mean(Y3))^2)
R_squared3 = 1 - RSS3 / TSS3

Y = c(Y1, Y2, Y3)
X3 = rbind(X, X, X)

beta = solve(t(X3) %*% X3) %*% t(X3) %*% Y
Y_pred = X3 %*% beta
residuals = Y - Y_pred

RSS = sum((Y - Y_pred)^2)
TSS = sum((Y - mean(Y))^2)
R_squared = 1 - RSS / TSS
```

Here, we display the value of the coefficients for the first few predictors in table 1. (See full set of coefficients in Appendix. (coming))

\begin{table}[h]
    \centering
    \begin{tabular}{l r r r r}
        \toprule
        Predictor & Model 1 & Model 2 & Model 3 & Model 4 \\
        \midrule
        Intercept  &  0.0057000007 &  0.0031776832 &  0.0026048919 &  0.0038275253 \\
        Neuron\_1  & -0.0007161198 &  0.0393005011 &  0.0353755239 &  0.0246533017 \\
        Neuron\_2  &  0.0277343380 & -0.0094288755 &  0.0107636538 &  0.0096897054 \\
        Neuron\_3  &  0.0016065925 &  0.0087287020 & -0.0014463192 &  0.0029629918 \\
        Neuron\_4  &  0.0027178334 & -0.0084431688 & -0.0005627236 & -0.0020960197 \\
        Neuron\_5  & -0.0057370064 & -0.0062233763 & -0.0054332369 & -0.0057978732 \\
        Neuron\_6  &  0.0012776469 &  0.0077952762 &  0.0055248920 &  0.0048659383 \\
        Neuron\_7  & -0.0024053779 & -0.0004962535 &  0.0012933504 & -0.0005360937 \\
        Neuron\_8  &  0.0059968300 & -0.0051638573 & -0.0011811706 & -0.0001160659 \\
        Neuron\_9  &  0.0044756078 & -0.0047117552 &  0.0022120294 &  0.0006586274 \\
        \bottomrule
    \end{tabular}
    \caption{Combined predictor coefficients from Models 1–4}
    \label{tab:combined_coefficients}
\end{table}

Table 5 summarizes the model performance, measured by $R^2$: 

\begin{table}[h]
    \centering
    \begin{tabular}{l r}
        \toprule
        Model & $R^2$ \\
        \midrule
        Model 1 & 0.075 \\
        Model 2 & 0.116 \\
        Model 3 & 0.153 \\
        Model 4 & 0.039 \\
        \bottomrule
    \end{tabular}
    \caption{$R^2$ values for different models}
\end{table}


\newpage
## 3.3 Comparing the Frequentist Models

After fitting the models, the residual plots shown in Figure 7 were created to compare the strength of our frequentist models. We also calculated the $R^2$ values for all of the models for strength comparison. None of the residual plots appear to be traditionally desirable, with all of them lacking a random spread across the x-axis and instead making similar patterns to one another. However, this does not mean the models are all bad or useless. Due to lack of certainty when making predictions in the neuroscience field, what might appear as a bad or weak model in a different statistical context is considered an acceptable model in a neuroscience context. The residual plots for our first frequentist model (the residual plots belonging to Neurons 234, 168, and 281) are slightly more desirable than the residual plot for our second frequentist model (the residual plot belonging to the Concatenated Model) as the spread across the x-axis seems to be more balanced and evenly split in our residual plots for our first frequentist models than our second. Additionally, the $R^2$ values of our first frequentist model (0.075, 0.116, and 0.153) are better than the $R^2$ value of 0.039 for our second frequentist model. Between these two details, we believe that considering each neuron separately and predicting for each one individually is better than trying to compile the predictions into a single model. This not only confirms our earlier hypothesis that Model 1 is the stronger model but also makes logical sense as each neuron connects to the overall neural network differently and so they will each be affected uniquely by every predictor neuron. Additionally, we can see in Figure 8 that the coefficients between the concatenated model and the separated models tend to differ, which makes sense considering what has been discovered about the $R^2$ values and the residual plots. The concatenated coefficients tend to be different from all of the separate models and worse at predicting, overall.

```{r figs7, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, fig.cap="\\label{fig:figs7}Residual Plots for Frequentist Models"}
par(mfrow=c(2, 2))

plot(Y1_pred, residuals1,
     xlab="Fitted Values (Neuron 234)",
     ylab="Residuals (Neuron 234)",
     main="Residuals Plot (Neuron 234)",
     col="yellow")
abline(h=0, col="red", lwd=2)

plot(Y2_pred, residuals2,
     xlab="Fitted Values (Neuron 168)",
     ylab="Residuals (Neuron 168)",
     main="Residuals Plot (Neuron 168)",
     col="blue")
abline(h=0, col="red", lwd=2)

plot(Y3_pred, residuals3,
     xlab="Fitted Values (Neuron 281)",
     ylab="Residuals (Neuron 281)",
     main="Residuals Plot (Neuron 281)",
     col="green")
abline(h=0, col="red", lwd=2)

plot(Y_pred, residuals,
     xlab="Fitted Values (Concatenated Model)",
     ylab="Residuals (Concatenated Model)",
     main="Residuals Plot (Concatenated Model)",
     col="purple")
abline(h=0, col="red", lwd=2)

par(mfrow=c(1,1))
```

```{r figs8, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=10, fig.cap="\\label{fig:figs8}Histogram of Difference in Coefficients between Frequentist Models"}
beta1dif = beta - beta1
beta2dif = beta - beta2
beta3dif = beta - beta3
par(mfrow=c(3,1))
hist(beta1dif, 
     main = "Difference between Neuron 234 and Concatenated Model", 
     xlab = "Difference in Coefficient", 
     col = "skyblue", 
     border = "black")

hist(beta2dif, 
     main = "Difference between Neuron 168 and Concatenated Model", 
     xlab = "Difference in Coefficient", 
     col = "skyblue", 
     border = "black")

hist(beta3dif, 
     main = "Difference between Neuron 281 and Concatenated Model", 
     xlab = "Difference in Coefficient", 
     col = "skyblue", 
     border = "black")
par(mfrow=c(1,1))
```

\newpage
# 4. Bayesian Analysis

Propose the Bayesian analysis you will work on during the rest of the quarter in this session.

## 4.1 Proposed Bayesian Model(s)

In this section,

formulate the Bayesian model(s) you are going to use to analyze your dataset. Be sure to first define the notations involved in the model(s).

discuss how you will elicit the prior(s).

## 4.2 Fitting the Bayesian model(s)

Propose how you will fit the proposed Bayesian models.

Propose how you will perform sensitivity analysis of the Bayesian models, i.e., how the posterior distribution is affected by the prior

Propose how you will check the MCMC convergence.

## 4.3 Prediction

In this section, propose how you can make predictions using the Bayesian model.

# 5. Discussion

In this section, discuss how you can improve your model.

# 6. Contributions

In this section, discuss the percentage of your contributions to the development final project proposal. Report the number of hours you have worked on the proposal, and the sections you are involved.

Please also discuss briefly the contributions of your teammate(s), as well as the help and support you got from your teammates(s).

# References

# Appendix

R code for Exploratory Data Analysis graphs:

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Neuron 234 Time Series Plot
spks_t = as.data.frame(t(spks))
neurons = paste("Neuron", 1:315, sep = "_")
colnames(spks_t) = neurons

time_points = seq(0, by = 30 / 1000, length.out = 148468)
spks_t$time = time_points

ggplot(spks_t, aes(x = time, y = Neuron_234)) +
  geom_line(size = 1) +
  labs(title = "Time Series", x = "Time", y = "Neuron 234") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Neuron Type Distribution
colnames(redlabels) = c("neuron_type")
redlabels$neuron_type = factor(redlabels$neuron_type, levels= c(TRUE, FALSE), labels=c("Inhibitory", "Excitatory"))
ggplot(redlabels, aes(x = neuron_type)) +
  geom_bar() +
  theme_minimal() +
  labs(title= "Number of Excitatory and Inhibitory Neurons", x = "Neuron Type", y = "Number of Neurons") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Pearson Correlation Heatmap
spks_t = as.data.frame(t(spks))
cor_matrix = cor(spks_t, use = "pairwise.complete.obs")
diag(cor_matrix) = NA

library(pheatmap)
pheatmap(cor_matrix, 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         main = "Pearson Correlation Heatmap",
         treeheight_row = 0,
         treeheight_col = 0,
         show_rownames = FALSE,
         show_colnames = FALSE
)
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Spatial Position Plot
colnames(cell_pos) = c("x", "y")
color_pos = cbind(cell_pos, redlabels)
ggplot(color_pos, aes(x = x, y = y, color = as.factor(neuron_type))) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("blue", "red")) +
  labs(title="Spatial Position of Neurons", color = "Neuron Type") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Distribution of Reach Times
colnames(calcium_reach_times) = c("start", "end")
calcium_reach_times$reach_time = calcium_reach_times$end - calcium_reach_times$start
ggplot(calcium_reach_times, aes(x=reach_time)) +
  geom_histogram(binwidth=20) +
  theme_minimal() +
  labs(title= "Distribution of Reach Times", x = "Reach Time", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Histogram of R^2 Values
load("r_squared_list.RData")
r_squared_vector <- unlist(r_squared_list)

hist(r_squared_vector, 
     main = "Histogram of R-squared values", 
     xlab = "R-squared", 
     col = "skyblue", 
     border = "black")
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Fitting Frequentist Models and Calculating their R^2
spks_t = as.data.frame(t(spks))
neurons = paste("Neuron", 1:315, sep = "_")
colnames(spks_t) = neurons

response_neurons = c("Neuron_234", "Neuron_168", "Neuron_281")
predictor_neurons = setdiff(neurons, response_neurons)
X = as.matrix(spks_t[, predictor_neurons])
X = cbind(Intercept = 1, X)

Y1 = spks_t[, "Neuron_234"]

beta1 = solve(t(X) %*% X) %*% t(X) %*% Y1
Y1_pred = X %*% beta1
residuals1 = Y1 - Y1_pred

RSS1 = sum((Y1 - Y1_pred)^2)
TSS1 = sum((Y1 - mean(Y1))^2)
R_squared1 = 1 - RSS1 / TSS1

Y2 = spks_t[, "Neuron_168"]

beta2 = solve(t(X) %*% X) %*% t(X) %*% Y2
Y2_pred = X %*% beta2
residuals2 = Y2 - Y2_pred

RSS2 = sum((Y2 - Y2_pred)^2)
TSS2 = sum((Y2 - mean(Y2))^2)
R_squared2 = 1 - RSS2 / TSS2

Y3 = spks_t[, "Neuron_281"]

beta3 = solve(t(X) %*% X) %*% t(X) %*% Y3
Y3_pred = X %*% beta3
residuals3 = Y3 - Y3_pred

RSS3 = sum((Y3 - Y3_pred)^2)
TSS3 = sum((Y3 - mean(Y3))^2)
R_squared3 = 1 - RSS3 / TSS3

Y = as.matrix(spks_t[, c("Neuron_234", "Neuron_168", "Neuron_281")])

beta = solve(t(X) %*% X) %*% t(X) %*% Y
Y_pred = X %*% beta
residuals = Y - Y_pred

RSS = sum((Y - Y_pred)^2)
TSS = sum((Y - mean(Y))^2)
R_squared = 1 - RSS / TSS
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Residual Plots for Frequentist Models
par(mfrow=c(2, 2))

plot(Y1_pred, residuals1,
     xlab="Fitted Values (Model 1)",
     ylab="Residuals (Model 1)",
     main="Residuals Plot (Model 1)",
     col="yellow")
abline(h=0, col="red", lwd=2)

plot(Y2_pred, residuals2,
     xlab="Fitted Values (Model 2)",
     ylab="Residuals (Model 2)",
     main="Residuals Plot (Model 2)",
     col="blue")
abline(h=0, col="red", lwd=2)

plot(Y3_pred, residuals3,
     xlab="Fitted Values (Model 3)",
     ylab="Residuals (Model 3)",
     main="Residuals Plot (Model 3)",
     col="green")
abline(h=0, col="red", lwd=2)

plot(Y_pred, residuals,
     xlab="Fitted Values (Model 4)",
     ylab="Residuals (Model 4)",
     main="Residuals Plot (Model 4)",
     col="purple")
abline(h=0, col="red", lwd=2)

par(mfrow=c(1,1))
```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
#Histograms showing difference between coefficients
beta1dif = beta - beta1
beta2dif = beta - beta2
beta3dif = beta - beta3
par(mfrow=c(3,1))
hist(beta1dif, 
     main = "Difference between Neuron 234 and Concatenated Model", 
     xlab = "Difference in Coefficient", 
     col = "skyblue", 
     border = "black")

hist(beta2dif, 
     main = "Difference between Neuron 168 and Concatenated Model", 
     xlab = "Difference in Coefficient", 
     col = "skyblue", 
     border = "black")

hist(beta3dif, 
     main = "Difference between Neuron 281 and Concatenated Model", 
     xlab = "Difference in Coefficient", 
     col = "skyblue", 
     border = "black")
par(mfrow=c(1,1))
```

